---
title: "feature_space"
author: "Shawn Schneidereit"
date: "`r Sys.Date()`"
output: 
  html_document:
   fig_width: 10
   fig_height: 6
   toc: TRUE
   toc_float: TRUE
   toc_collapsed: FALSE
   code_folding: hide
   theme: united
   highlight: tango


---

```{r, cache=TRUE, setup, include=FALSE}
# library
library(terra)     # raster data maipulations
#library(sf)
library(viridis)   # color palletes
library(tidyverse) # data manipulation
library(patchwork) # for grid visualization
library(ggridges)  # ridge plots
library(ggpmisc)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```


# Data import 

```{r, cache=TRUE, include=FALSE}
# data import and wrangel 

raw_fs <- read.csv("data/feature_space/20220602_FS_LS4_LS5_LS7_LS8_LS9_1985_2022.csv")

```

## QA band filtering

### Read QA band

```{r, cache=TRUE, include=FALSE}
### What are the three most frequent values in the qa band?
raw_fs %>% group_by(QAI) %>% tally() %>% arrange(-n)

# can be used in cross reference with https://force-eo.readthedocs.io/en/latest/howto/qai.html?highlight=qai%20bits
rev(as.numeric(intToBits(8192)[1:16]))  
rev(as.numeric(intToBits(0)[1:16])) 
rev(as.numeric(intToBits(10240)[1:16]))  
```

### Filtering function

```{r, cache=TRUE, echo=TRUE}
# high confidence clouds or high confidence cloud shadows or fill values
mask_medconf <- function(x){
  # use base R function to convert numeric into bits
  bs <- intToBits(x)
  if ( ((bs[1]) | # cloud
       # (bs[2]) | # low cloud confidence
        (bs[3]) | # shadow
        (bs[4]) | # snow
        (bs[5]) | # water
        (bs[6]) | # aerosol (only low quality)
        (bs[8]) | # subzero
        (bs[9]) | # saturation
        (bs[10]) | # illumination
        (bs[11]) | # saturation
        (bs[13]) | # slope
        (bs[14])) # vapor
       == 1){
    return("flag") } else {
      return("valid")
    }
}



# due to the nature of the intToBits function it is not possible to perform 
# this step in a mutate pipe. Therefor a vector is created via lapply, 
# joined to then joined to the fs_LND df and finally flaged pixels can be removed
mask_idex <- lapply(raw_fs$QAI, mask_medconf) %>% unlist()

fs_LND_filtered <- raw_fs %>%
  mutate(QAI = mask_idex) %>% 
  filter(QAI == "valid")
```

Quite important to note here is that while the absolute number of pixels additionally filtered with more strict conditions applied is quite small, increasing these conditions did yield effective improvements in removing additional outlyers in the data (particularly evident in direct pixel scatter plot comparisons between LND05/07/08). Therefor the deliberate inclusions/exclusion of certain filtering criterion could disproportionately sway the final results of sensor comparisons and feature spaces. 


### Katjas method compairson

```{r, cache=TRUE, eval=TRUE}
# Katja´s method to filter QAI values
raw_fs_KK <- raw_fs

raw_fs_KK$qa_cloud <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 1), 3)
raw_fs_KK$qa_shadow <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 3), 1)
raw_fs_KK$qa_snow <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 4), 1)
raw_fs_KK$qa_water <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 5), 1)
raw_fs_KK$qa_aerosol <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 6), 3)
raw_fs_KK$qa_subzero <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 8), 1)
raw_fs_KK$qa_saturation <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 9), 1)
raw_fs_KK$qa_zenith <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 10), 1)
raw_fs_KK$qa_illumination <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 11), 3)
raw_fs_KK$qa_slope <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 13), 1)
raw_fs_KK$qa_vapor <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 14), 1)


raw_fs_KK$quality <- 0
raw_fs_KK$quality[raw_fs_KK$qa_cloud != 0 |raw_fs_KK$qa_snow != 0 |raw_fs_KK$qa_shadow != 0] <- 1
raw_fs_KK <- raw_fs_KK[raw_fs_KK$quality == 0, ]
```


```{r, cache=TRUE, eval=TRUE}
print(paste("number of observations following SHS method for QAI filtering:", round(nrow(fs_LND_filtered)/6)) )
print(paste("number of observations following KK method QAI filtering:", round(nrow(raw_fs_KK)/6)) )
print(paste("The difference in included observation:", round(nrow(fs_LND_filtered)/nrow(raw_fs_KK), digits = 2), "% (n=",  (nrow(fs_LND_filtered)-nrow(raw_fs_KK)),  ")") )
  


```

As seen above the differences in the filtering methods and parameterization is very marginal, at ~1% difference in amount of included observations

##Additional data filtering and auxilirary varaible and indicity computation

```{r, cache=TRUE}

fs_LND <- fs_LND_filtered %>% 
  # reduce file size for testing
  #sample_n(10000) %>% 
  # filter out extreme values
  filter(across(BLUE:SWIR2, ~ . > 0),
         across(BLUE:SWIR2, ~ . < 10000)) %>% 
  mutate(date   = as.Date(str_sub(scene, 1, 9),format = "%Y%m%d"),
         month  = lubridate::month(date),
         year   = lubridate::year(date), 
         month_year = paste0(month,"_",year),
         doy    = lubridate::yday(date),
         sensor = str_sub(scene, -5, -1),
         SWIR_ratio = SWIR2/SWIR1,
         NDVI   = ((NIR-RED)/(NIR+RED)),
         NDTI   = ((SWIR1-SWIR2)/(SWIR1+SWIR2))) # source: https://www.mdpi.com/2072-4292/10/10/1657/htm
  
```


# Spectra vis

## NDVI time series
```{r, cache=TRUE}
ggplot(fs_LND, aes(doy, NDVI, color=year, group=year)) +
  geom_smooth() +
  scale_colour_viridis_c(option = "D") +
  theme_minimal()
```

Here it can be seen that over time, average NDVI maximum values tend to increase over time, particularly early in the year before the growing season. Over such a long time frame it is possible though that this chance could be attributed to more irrigation, climate change, or LULC rather been strictly being resultant of deviations in sensor spec?

## NDTI time series
```{r, cache=TRUE}

ggplot(na.omit(fs_LND), aes(doy, NDTI, color=year, group=year)) +
  geom_smooth() +
  scale_colour_viridis_c(option = "D") +
  theme_minimal()
```

Similar story here, with NDTI decreasing over time. NDTI is also negatively associated with tillage, so the dip in the growing season is explained by more bare soil (tilled) locations being progressively covered by PV. Sudden decreases during the growing season indicate a tilling event, where as a slow decline at the end of the growing season are more likely to indicate PV senescence. 

## Data wrangel into long

```{r, cache=TRUE}
fs_LND_long <- fs_LND %>% 
  pivot_longer(cols=c("BLUE":"SWIR2"),names_to = "wavelength", values_to = "reflectance") %>% 
  as.data.frame()  %>% 
    mutate(wavelength_num = case_when(wavelength == "BLUE" ~ 482,
                                      wavelength == "GREEN" ~ 562,
                                      wavelength == "RED" ~ 655,
                                      wavelength == "NIR" ~ 865,
                                      wavelength == "SWIR1" ~ 1610,
                                      wavelength == "SWIR2" ~ 2200),
            reflectance = reflectance/100) # scale to percent
```


## Simple mean reflectance over time
```{r, cache=TRUE}
ggplot(fs_LND_long, aes(wavelength_num, reflectance, color=year, group=year)) +
   # geom_smooth(formula = y ~ s(x, bs = "cs", k=6)) +
   stat_summary(fun=mean, geom="line", size = 1) + # draw a mean line in the data
  scale_colour_viridis_c(option = "D") +
  theme_minimal()

```

When plotting a simple mean of spectra across years it can be seen that more recent years dont defaco have higher reflectance values across the entire electromagnetic spectrum


## Boxplot

```{r, cache=TRUE}
ggplot(fs_LND_long, aes(wavelength, reflectance, color=sensor)) +
 # geom_jitter() +
  geom_boxplot() +
  scale_colour_viridis_d(option = "D") +
  theme_minimal()
```

Looks like lots of extreme values (even post QAI filter), *but* when looking at the next density plot section, you can that relatively speaking, these extreme outlying values are exceedinly rare and marginal.

## Density

```{r, cache=TRUE}


# facet wrap by wavelength
ggplot(fs_LND_long, aes(reflectance, color=sensor, fill=sensor)) +
  geom_density(alpha = 0.05) +
  #geom_jitter() +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
  theme_minimal() +
  guides(col = guide_legend(nrow = 3))+
  theme(legend.position = "bottom")  +
  facet_wrap(~wavelength)


# facet wrap by sensor
ggplot(fs_LND_long, aes(reflectance, color=wavelength, color=wavelength)) +
  geom_density(alpha = 0.05) +
  #geom_jitter() +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
  theme_minimal() +
  guides(col = guide_legend(nrow = 3))+
  theme(legend.position = "bottom")  +
  facet_wrap(~sensor)


```

## Ridges

```{r, cache=TRUE}
ggplot(fs_LND_long, aes(x = reflectance, y = sensor, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "reflectance", option = "D") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 75)) +
  theme_minimal() +
  facet_wrap(~wavelength)

ggplot(fs_LND_long, aes(x = reflectance, y = wavelength, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "reflectance", option = "D") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 75)) +
  theme_minimal() +
  facet_wrap(~sensor)
  
```


# Feature space vis

## Import reference data

```{r, cache=TRUE}
reference_spectra <- read.csv("data/feature_space/sli_gen_dark_soils_0p4.csv",
                              encoding = "UTF-8") %>% 
  mutate(SWIR_ratio = SWIR2/SWIR1,
         NDVI   = ((NIR-RED)/(NIR+RED)),
         NDTI   = ((SWIR1-SWIR2)/(SWIR1+SWIR2)))

ggplot(reference_spectra, aes(NDVI, SWIR_ratio, color = cover)) +
  geom_point() +
  scale_colour_viridis_d(option = "D") +
  theme_minimal() 
```

## Plot complete data

```{r, cache=TRUE}
# No dissagregation
ggplot(fs_LND, aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 300) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2))
```

Here we can see that there are two dense zones in the spectral feature spaces. The smaller and fainter zone is predominately populated by landsat 7 pixels (clearly seen in the following plot)

## Facet by sensor
```{r, cache=TRUE}
ggplot(fs_LND, aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 200) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~sensor)
```

Here we can see that the spectral library fits particularly well to Landsat 8 (probably do to the fact that is was created using lansat 8 pixels?). 

## Facet by month
```{r, cache=TRUE}
ggplot(fs_LND, aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 100) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~month)
```

Here we can also see that across all sensors, the spectral library seems to triangulate the data points better during growing season months. 

## Facet by year

_Displayed with only point per reference class to improve interpretability_ 

### 1985-2000
```{r, cache=TRUE, warning=FALSE}
# FROM 1985-2000
fs_LND |> filter(year<2001) %>%
ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 100) +
  geom_point(data=reference_spectra[c(1,3,25),], aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)


```

### 2001-2022
```{r, cache=TRUE, warning=FALSE}
# FROM 2001-2022
fs_LND |> filter(year>2001) %>%
ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 100) +
  geom_point(data=reference_spectra[c(1,3,25),], aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)

```

## Feature space by sensor and year

### Summerarised boxplot overview

```{r, cache=TRUE}
ggplot(fs_LND, aes(x=, y=NDVI/SWIR_ratio, group=year, fill=sensor, color=sensor)) +
  geom_boxplot(alpha=0.3) +
 # geom_jitter(alpha=0.3) +
 #geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  theme_minimal() +
#  scale_x_continuous(breaks=seq(1,12,1)) +
 # scale_y_continuous(limits = c(0, 1000)) +
  facet_wrap(~sensor, scales = "free_x")
```

Here is a simple boxplot overview per sensor over time. To aggregate the the NVDI and SWIR-ratio of the regular feature space scatter plot a simple ratio between NDVI and SWIR-ratio was taken (NDVI/SWIR_ratio). Overall the mean ratio and distribution of values seems to be relatively consistent, with the exceptions of some occasional outlines. The years 2003 and 2018 show exceptionally large deviations from the rest of the timeseries. Yet when comparing these years between different sensors it can be seen that this dip is consistent for multiple sensors, potentially indicating that this deviation yearly weather (etc) related phenomenon and not an artifact resulting from sensor instability?

_The following plots contain feature space scatter plots, split by year and sensor. The dotted black line represents the overall regression computed across all sensors and years combined and should provide a base line to check years/sensor combinations scatter plots against._

### Landsat 4
```{r, cache=TRUE}
my.formula <- y ~ x

fs_LND |> filter(sensor =="LND04") %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
 stat_bin_hex(bins = 50) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```




### Landsat 5 1985:1995


```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND05",
                 year %in% c(1985:1995)) %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 50) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```

### Landsat 5 1995:2005
```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND05",
                 year %in% c(1995:2005)) %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 40) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```

### Landsat 5 2005:2011
```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND05",
                 year %in% c(2005:2011)) %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 40) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```


### Landsat 7 1999:2009
```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND07",
                 year %in% c(1999:2009)) %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 50) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```

### Landsat 7 2009:2022
```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND07",
                 year %in% c(2009:2022)) %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
 stat_bin_hex(bins = 50) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```




### Landsat 8
```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND08") %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 50) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```

### Landsat 9
```{r warning=FALSE, cache=TRUE}
fs_LND |> filter(sensor =="LND09") %>%
  ggplot(., aes(NDVI, SWIR_ratio)) +
  stat_bin_hex(bins = 40) +
  geom_abline(intercept=.78,slope=-0.5, linetype=2) +
  # stat_smooth(method = "lm", formula = my.formula, color="red") +
  # stat_poly_eq(formula = my.formula, 
  #              aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
  #              size = 2.5,
  #              parse = TRUE,
  #              label.x =1.2,
  #              label.y =1.05) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red",
             size=0.6, alpha=0.9) +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~year)
```


## NDTI feature space

```{r, cache=TRUE}
ggplot(fs_LND, aes(NDVI, NDTI)) +
  geom_point() +
  stat_bin_hex(bins = 250) +
  geom_point(data=reference_spectra, aes(NDVI, NDTI), color ="red") +
  scale_fill_continuous(type = "viridis") +
   scale_x_continuous(expand = c(0, 0), limits = c(-0.1, 1)) +
   scale_y_continuous(expand = c(0, 0), limits = c(-0.25, 0.8)) +
  theme_minimal()# +
#  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
#  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) 
```

## Facet by sensor
```{r, cache=TRUE}
ggplot(fs_LND, aes(NDVI, NDTI)) +
  stat_bin_hex(bins = 200) +
  geom_point(data=reference_spectra, aes(NDVI, NDTI), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-0.25, 1.2)) +
  facet_wrap(~sensor)
```

## Facet by month
```{r, cache=TRUE}
ggplot(fs_LND, aes(NDVI, NDTI)) +
  stat_bin_hex(bins = 100) +
  geom_point(data=reference_spectra, aes(NDVI, NDTI), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-0.25, 1.2)) +
  facet_wrap(~month)
```

```{r, cache=TRUE, include==FALSE}

# old computationally exorbinant method of and plotting density plots 

# # Get density of points in 2 dimensions.
# # @param x A numeric vector.
# # @param y A numeric vector.
# # @param n Create a square n by n grid to compute density.
# # @return The density within each square.
# get_density <- function(x, y, ...) {
#   dens <- MASS::kde2d(x, y, ...)
#   ix <- findInterval(x, dens$x)
#   iy <- findInterval(y, dens$y)
#   ii <- cbind(ix, iy)
#   return(dens$z[ii])
# }
# 
# fs_density <- as.data.frame(fs_LND) %>%
#   filter(SWIR_ratio != "Inf") %>% 
#   mutate(density = get_density((.)$NDVI, (.)$SWIR_ratio, n = 500)) 

# ggplot(fs_density, aes(NDVI, SWIR_ratio, color = density)) +
#   geom_point() +
#   geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
#   scale_colour_viridis_c(option = "D") +
#   theme_minimal() +
#   scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
#   facet_wrap(~sensor)

```

# single location band compairison

## Overview of observations over time by sensor


## Temporal overvire and year resolution
```{r, cache=TRUE}

fs_LND |>  
  group_by( year, sensor) %>% 
  summarise(n_observation = n()) %>%  
  
ggplot(., aes(year, n_observation, color=sensor)) +
  geom_line() +
  scale_fill_continuous(type = "viridis") +
  scale_colour_viridis_d(option = "D") +
  theme_minimal() 
```


## Temporal overview at month resolution
```{r, cache=TRUE}
fs_LND_obs_overview <- fs_LND %>% 
  group_by(month, year, sensor) %>% 
  summarise(n_observation = n()) %>%
  mutate(timestamp = zoo::as.yearmon(paste0(month,"_", year), format="%m_%Y"))
```


## Monthly time series
```{r, cache=TRUE}

ggplot(fs_LND_obs_overview, aes(timestamp, n_observation, color=sensor)) +
  geom_line(size=1) +
  #geom_smooth( ) +
  scale_fill_continuous(type = "viridis") +
  scale_colour_viridis_d(option = "D") +
  scale_x_continuous(breaks=seq(1984,2022,2)) +
  theme_minimal()

```


The acute fluctuations in observation numbers seems to stem from the fact that during winter months there is a lower observation density (plotted below)


## Observation density by month

_(same data with some different vis options)_

```{r, cache=TRUE, warning=FALSE}


ggplot(fs_LND_obs_overview, aes(x=month, y=n_observation, group=month,fill=sensor, color=sensor)) +
 # geom_boxplot(alpha=0.3, notch = T) +
  geom_violin(alpha=0.3) +
  geom_jitter(alpha=0.3) +
 #geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  theme_minimal() +
  scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(limits = c(0, 1000)) +
  facet_wrap(~sensor)


ggplot(fs_LND_obs_overview, aes(x=month, y=n_observation, group=month,fill=sensor, color=sensor)) +
  geom_boxplot(alpha=0.3) +
  geom_jitter(alpha=0.3) +
 #geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  theme_minimal() +
  scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(limits = c(0, 1000)) +
  facet_wrap(~sensor)

ggplot(fs_LND_obs_overview, aes(x=month, y=n_observation, group=month,fill=sensor, color=n_observation)) +
 geom_boxplot(alpha=0.5) +
 # geom_violin(alpha=0.3) +
  geom_jitter(alpha=0.9) +
 #geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01) +
  scale_colour_viridis_c(option = "D") +
  scale_fill_viridis_d(option = "D") +
  theme_minimal() +
  scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(limits = c(0, 1000)) +
  facet_wrap(~sensor)


```

The density of observations in summer months are far greater (although more variable), While during winter months are seem to be constantly few observations. This trend is stable across sensors, so mostly likely can largely attributed to increased cloud/snow cover during winter months. 


##  *Exact* matches of location and *loose* match time of observation
```{r, cache=TRUE, cache=TRUE}
fs_LND_loose <- fs_LND_long %>% 
  # filter(sensor %in% c("LND07", "LND08")) %>% 
  mutate(obs_time_place = paste0(plyr::round_any(doy, 2, f = floor),"_", year,"_",POINT_ID)) %>% 
  group_by(obs_time_place) %>% 
  filter(n() > 6)


```

Total number of observations that match the loose condition of being within ten days of each other: `r (nrow(fs_LND_loose)/6) `

Days of year with available observations: `r sort(unique(fs_LND_loose$doy)) `

## Sample compairison of LND07 and LND08 spectra 

```{r, cache=TRUE}
for (i in 1:4) {
  
  scens <- unique(fs_LND_loose$obs_time_place)[c((((i-1)*20)+1):((i)*20))]
  
  fs_LND_loose_subset <- fs_LND_loose %>% 
    filter(obs_time_place %in% scens) |> 
    filter(sensor %in% c("LND07", "LND08")) 
  
  print(ggplot( fs_LND_loose_subset, aes(wavelength_num, reflectance, color=sensor, fill=sensor)) +
    # geom_density(alpha = 0.05) +
    geom_line() +
   # scale_colour_viridis_d(option = "D") +
    #scale_fill_viridis_d(option = "D") +
    scale_x_continuous(expand = c(0, 0)) +
    #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
    theme_minimal() +
    guides(col = guide_legend(nrow = 3))+
     theme(legend.position = "bottom", 
           axis.text.y = element_text(angle = 45),
           strip.text.y = element_text(size = 8, angle = 330))  +
    facet_wrap( ~obs_time_place) )
  }

```



# Scatterplots 




```{r, cache=TRUE}
fs_LND_scatter <- fs_LND_loose |> 
  select(obs_time_place,sensor,wavelength,reflectance) |> 
  pivot_wider(names_from = sensor, values_from = c(reflectance)) |> 
  rename(reflectance_LND04 = LND04,
         reflectance_LND05 = LND05,
         reflectance_LND07 = LND07,
         reflectance_LND08  = LND08,
         reflectance_LND09= LND09)
```


## Scatterplot of Landsat 7 and 8 specta

```{r, cache=TRUE}

my.formula <- y ~ x

fs_LND_scatter_LND_7_8 <- fs_LND_scatter |> select(-reflectance_LND04,-reflectance_LND05, -reflectance_LND09) |> 
  na.omit()

ggplot(fs_LND_scatter_LND_7_8,aes(reflectance_LND07, reflectance_LND08)) +
  stat_bin_hex(bins = 100) + 
  stat_smooth(method = "lm", formula = my.formula, color="red") +
  stat_poly_eq(formula = my.formula, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE,
               label.x =1.2,
               label.y =1.05) +
  geom_abline(intercept=1,slope=1, linetype=2) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  annotate("text", x=7, y=105,
           label= paste0("R2 =",round((cor((fs_LND_scatter_LND_7_8$reflectance_LND07), 
                                           (fs_LND_scatter_LND_7_8$reflectance_LND08),
                                           use="complete.obs")^2),2)), size=4.5, hjust=0) + 
  
  annotate("text", x= 7, y=100, hjust=0, size=4.5,
           label= paste0("RMSE ==",round(sqrt(mean(((fs_LND_scatter_LND_7_8$reflectance_LND07)-
                                                      (fs_LND_scatter_LND_7_8$reflectance_LND08))^2,
                                                   na.rm=TRUE)),2)), parse=TRUE)+
  
  annotate("text", x=7, y= 95, size=4.5, hjust=0,
           label=paste0("Bias = ", round((mean((fs_LND_scatter_LND_7_8$reflectance_LND07), na.rm=TRUE) -
                                            mean((fs_LND_scatter_LND_7_8$reflectance_LND08),na.rm=TRUE)),2)))+
  
  annotate("text", x=7, y=90,size=4.5,hjust=0, 
           label=paste0("MAE = ", round(mean(abs((fs_LND_scatter_LND_7_8$reflectance_LND07)-                                          
                                                   (fs_LND_scatter_LND_7_8$reflectance_LND08))),2)))
```

## Compairson by wavelength

```{r, cache=TRUE}
ggplot(fs_LND_scatter, aes(reflectance_LND07, reflectance_LND08)) +
  stat_bin_hex(bins = 100) +
  stat_smooth(method = "lm", formula = my.formula, color="red") +
  stat_poly_eq(formula = my.formula, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE,
               label.x =1.2,
               label.y =1.05) +
  geom_abline(intercept=1,slope=1, linetype=2) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  facet_wrap(~wavelength, 
             scales="free"
  ) 

kable(
fs_LND_scatter |> 
  select(-reflectance_LND04,-reflectance_LND05,-reflectance_LND09) |> 
  group_by(wavelength) |> 
  summarise(R2   = round((cor((reflectance_LND07), 
                            (reflectance_LND08),
                            use="complete.obs")^2),2),
            
            RMSE = round(sqrt(mean(((reflectance_LND07)-
                                      (reflectance_LND08))^2, na.rm=TRUE)),2),
            
            Bias = round((mean((reflectance_LND07), na.rm=TRUE) -
                            mean((reflectance_LND08), na.rm=TRUE)),2),
            
            MAE = round(mean(abs((reflectance_LND07)-
                                   (reflectance_LND08)), na.rm=TRUE),2))
) %>%
  kable_styling(full_width = T)


```



# Scatterplot of Landsat 5 and 7 specta

```{r, cache=TRUE}

fs_LND_scatter_LND_5_7 <- fs_LND_scatter |> select(-reflectance_LND04,-reflectance_LND08, -reflectance_LND09) |> 
  na.omit()


ggplot(fs_LND_scatter_LND_5_7, aes(reflectance_LND07, reflectance_LND05)) +
stat_bin_hex(bins = 100) +
  geom_abline(intercept=1,slope=1, linetype=2) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() + 
  stat_smooth(method = "lm", formula = my.formula, color="red") +
  stat_poly_eq(formula = my.formula, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE,
               label.x =1.2,
               label.y =1.05) +
  annotate("text", x=7, y=105,
           label= paste0("R2 =",round((cor((fs_LND_scatter_LND_5_7$reflectance_LND07), 
                                           (fs_LND_scatter_LND_5_7$reflectance_LND05),
                                           use="complete.obs")^2),2)), size=4.5, hjust=0) + 
  
  annotate("text", x= 7, y=100, hjust=0, size=4.5,
           label= paste0("RMSE ==",round(sqrt(mean(((fs_LND_scatter_LND_5_7$reflectance_LND07)-
                                                    (fs_LND_scatter_LND_5_7$reflectance_LND05))^2,
                                                   na.rm=TRUE)),2)), parse=TRUE)+
  
  annotate("text", x=7, y= 95, size=4.5, hjust=0,
           label=paste0("Bias = ", round((mean((fs_LND_scatter_LND_5_7$reflectance_LND07), na.rm=TRUE) -
                                          mean((fs_LND_scatter_LND_5_7$reflectance_LND05),na.rm=TRUE)),2)))+
  
  annotate("text", x=7, y=90,size=4.5,hjust=0, 
           label=paste0("MAE = ", round(mean(abs((fs_LND_scatter_LND_5_7$reflectance_LND07)-
                                                (fs_LND_scatter_LND_5_7$reflectance_LND05))),2))) 

```

## Compairson by wavelength

```{r, cache=TRUE}
ggplot(fs_LND_scatter, aes(reflectance_LND07, reflectance_LND05)) +
  stat_bin_hex(bins = 100) +
  stat_smooth(method = "lm", formula = my.formula, color="red") +
  stat_poly_eq(formula = my.formula, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE,
               label.x =1.2,
               label.y =1.05) +
  geom_abline(intercept=1,slope=1, linetype=2) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  facet_wrap(~wavelength, 
             scales="free"
  ) 

kable(
fs_LND_scatter |> 
  select(-reflectance_LND04,-reflectance_LND08,-reflectance_LND09) |> 
  group_by(wavelength) |> 
  summarise(R2   = round((cor((reflectance_LND07), 
                            (reflectance_LND05),
                            use="complete.obs")^2),2),
            
            RMSE = round(sqrt(mean(((reflectance_LND07)-
                                      (reflectance_LND05))^2, na.rm=TRUE)),2),
            
            Bias = round((mean((reflectance_LND07), na.rm=TRUE) -
                            mean((reflectance_LND05), na.rm=TRUE)),2),
            
            MAE = round(mean(abs((reflectance_LND07)-
                                   (reflectance_LND05)), na.rm=TRUE),2))
) %>%
  kable_styling(full_width = T)


```


Overall the correspondence between both Landsat 5/7 & 7/8 seems to be pretty good? R^2 values are around 90% and the MAE around a tick above 2%, which doesn't seem to drastic? In both comparison cases LND07 seemed to have (very) slightly lower reluctance values. Notable is the occasional occurrence of extreme outlyers across all the visable wavelengths in the spectra. Given that polluted pixels such as snow and clouds should have been removed through the QAI filters I am unsure what could be the source of these observed deviations (especially since they are not consistent across wavelengths). 

Some float from the ideal 1-to-1 line can be likely explained by the two day buffer in doy matching between pixels. The relative infrequence of these extreme outlyers could be explained by a major LULC event such as mowing or harvesting occurring exactly in this potential two day difference in observation date? Yet this does not then adress why the large deviance in reflectance would only be observed for blue,green,red pixels and not NIR+ ones...



























