---
title: "feature_space"
author: "Shawn Schneidereit"
date: "`r Sys.Date()`"
output: 
  html_document:
   fig_width: 10
   fig_height: 6
   toc: TRUE
   toc_float: TRUE
   toc_collapsed: FALSE
   code_folding: hide
   theme: united
   highlight: tango


---

```{r, cache=TRUE, setup, include=FALSE}
# library
library(terra)
#library(sf)
library(viridis)
library(tidyverse)
library(patchwork)
library(ggridges)
knitr::opts_chunk$set(echo = TRUE)
```


# Data import 

```{r, cache=TRUE, include=FALSE}
# data import and wrangel 

raw_fs <- read.csv("data/feature_space/20220602_FS_LS4_LS5_LS7_LS8_LS9_1985_2022.csv")

```

## QA band filtering

### Read QA band

```{r, cache=TRUE, include=FALSE}
### What are the three most frequent values in the qa band?
raw_fs %>% group_by(QAI) %>% tally() %>% arrange(-n)

# can be used in cross reference with https://force-eo.readthedocs.io/en/latest/howto/qai.html?highlight=qai%20bits
rev(as.numeric(intToBits(8192)[1:16]))  
rev(as.numeric(intToBits(0)[1:16])) 
rev(as.numeric(intToBits(10240)[1:16]))  
```

### filtering function

```{r, cache=TRUE}
# high confidence clouds or high confidence cloud shadows or fill values
mask_medconf <- function(x){
  bs <- intToBits(x)
  if ( ((bs[1]) | # cloud
        (bs[3]) | # shadow
        (bs[4]) | # snow
        (bs[5]) | # water
        (bs[6]) | # aerosol (only low quality)
        (bs[8]) | # subzero
        (bs[9]) | # saturation
        (bs[10])) # High sun zenith flag
       == 1){
    return("flag") } else {
      return("valid")
    }
}



# due to the nature of the intToBits function it is not possible to perform 
# this step in a mutate pipe. Therefor a vector is created via lapply, 
# joined to then joined to the fs_LND df and finally flaged pixels can be removed
mask_idex <- lapply(raw_fs$QAI, mask_medconf) %>% unlist()

fs_LND_filtered <- raw_fs %>%
  mutate(QAI = mask_idex) %>% 
  filter(QAI == "valid")
```

### Katjas method compairson

```{r, cache=TRUE, eval=TRUE}
# Katja´s method to filter QAI values
raw_fs_KK <- raw_fs

raw_fs_KK$qa_cloud <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 1), 3)
raw_fs_KK$qa_shadow <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 3), 1)
raw_fs_KK$qa_snow <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 4), 1)
raw_fs_KK$qa_water <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 5), 1)
raw_fs_KK$qa_aerosol <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 6), 3)
raw_fs_KK$qa_subzero <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 8), 1)
raw_fs_KK$qa_saturation <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 9), 1)
raw_fs_KK$qa_zenith <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 10), 1)
raw_fs_KK$qa_illumination <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 11), 3)
raw_fs_KK$qa_slope <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 13), 1)
raw_fs_KK$qa_vapor <- bitwAnd(bitwShiftR(raw_fs_KK$QAI, 14), 1)


raw_fs_KK$quality <- 0
raw_fs_KK$quality[raw_fs_KK$qa_cloud != 0 |raw_fs_KK$qa_snow != 0 |raw_fs_KK$qa_shadow != 0] <- 1
raw_fs_KK <- raw_fs_KK[raw_fs_KK$quality == 0, ]
```


```{r, cache=TRUE, eval=TRUE}
print(paste("number of observations following SHS method for QAI filtering:", (nrow(fs_LND_filtered)/6)) )
print(paste("number of observations following KK method QAI filtering:", (nrow(raw_fs_KK)/6)) )
print(paste("The difference in included observation:", (nrow(fs_LND_filtered)/nrow(raw_fs_KK)), "% (n=",  (nrow(fs_LND_filtered)-nrow(raw_fs_KK)),  ")") )
  


```

As seen above the differences in the filtering methods and parameterization is very marginal, at ~1% difference in amount of included observations

##Additional data filtering and auxilirary varaible and indicity computation

```{r, cache=TRUE}

fs_LND <- fs_LND_filtered %>% 
  # reduce file size for testing
  #sample_n(10000) %>% 
  # filter out extreme values
  filter(across(BLUE:SWIR2, ~ . > 0),
         across(BLUE:SWIR2, ~ . < 10000)) %>% 
  mutate(date   = as.Date(str_sub(scene, 1, 9),format = "%Y%m%d"),
         month  = lubridate::month(date),
         year   = lubridate::year(date), 
         month_year = paste0(month,"_",year),
         doy    = lubridate::yday(date),
         sensor = str_sub(scene, -5, -1),
         SWIR_ratio = SWIR2/SWIR1,
         NDVI   = ((NIR-RED)/(NIR+RED)),
         NDTI   = ((RED-SWIR2)/(RED+SWIR2))) # source: https://www.mdpi.com/2072-4292/10/10/1657/htm
  
```

# spectra vis

## NDVI time series
```{r, cache=TRUE}
ggplot(fs_LND, aes(doy, NDVI, color=year, group=year)) +
  geom_smooth() +
  scale_colour_viridis_c(option = "D") +
  theme_minimal()
```

Here it can be seen that over time, average NDVI maximum values tend to increase over time, particularly early in the year before the growing season. Over such a long time frame it is possible though that this chance could be attributed to more irrigation, climate change, or lULC rather been strictly being resultant of deviations in sensor spec?

## NDTI time series
```{r, cache=TRUE}

ggplot(fs_LND, aes(doy, NDTI, color=year, group=year)) +
  geom_smooth() +
  scale_colour_viridis_c(option = "D") +
  theme_minimal()
```

Simular story here, with NDTI decreasing over time

## data wrangel into long

```{r, cache=TRUE}
fs_LND_long <- fs_LND %>% 
  pivot_longer(cols=c("BLUE":"SWIR2"),names_to = "wavelength", values_to = "reflectance") %>% 
  as.data.frame()  %>% 
    mutate(wavelength_num = case_when(wavelength == "BLUE" ~ 482,
                                      wavelength == "GREEN" ~ 562,
                                      wavelength == "RED" ~ 655,
                                      wavelength == "NIR" ~ 865,
                                      wavelength == "SWIR1" ~ 1610,
                                      wavelength == "SWIR2" ~ 2200))
```


## simple mean reflectance over time
```{r, cache=TRUE}
ggplot(fs_LND_long, aes(wavelength_num, reflectance, color=year, group=year)) +
   # geom_smooth(formula = y ~ s(x, bs = "cs", k=6)) +
   stat_summary(fun=mean, geom="line", size = 1) + # draw a mean line in the data
  scale_colour_viridis_c(option = "D") +
  theme_minimal()

```

When plotting a simple mean of spectra across years it can be seen that more recent years dont defaco have higher reflectance values across the entire electromagnetic spectrum


## boxplot

```{r, cache=TRUE}
ggplot(fs_LND_long, aes(wavelength, reflectance, color=sensor)) +
 # geom_jitter() +
  geom_boxplot() +
  scale_colour_viridis_d(option = "D") +
  theme_minimal()
```

Looks like lots of extreme values (even post QAI filter), *but* when looking at the next density plot section, you can that relatively speaking, these extreme outlying values are exceedinly rare and marginal.

## density

```{r, cache=TRUE}


# facet wrap by wavelength
ggplot(fs_LND_long, aes(reflectance, color=sensor, fill=sensor)) +
  geom_density(alpha = 0.05) +
  #geom_jitter() +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
  theme_minimal() +
  guides(col = guide_legend(nrow = 3))+
  theme(legend.position = "bottom")  +
  facet_wrap(~wavelength)


# facet wrap by sensor
ggplot(fs_LND_long, aes(reflectance, color=wavelength, color=wavelength)) +
  geom_density(alpha = 0.05) +
  #geom_jitter() +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
  theme_minimal() +
  guides(col = guide_legend(nrow = 3))+
  theme(legend.position = "bottom")  +
  facet_wrap(~sensor)


```

## ridges

```{r, cache=TRUE}
ggplot(fs_LND_long, aes(x = reflectance, y = sensor, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "reflectance", option = "D") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 15000)) +
  theme_minimal() +
  facet_wrap(~wavelength)

ggplot(fs_LND_long, aes(x = reflectance, y = wavelength, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "reflectance", option = "D") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 15000)) +
  theme_minimal() +
  facet_wrap(~sensor)
  
```

# feature space vis

## import reference data

```{r, cache=TRUE}
reference_spectra <- read.csv("data/feature_space/sli_gen_dark_soils_0p4.csv",
                              encoding = "UTF-8") %>% 
  mutate(SWIR_ratio = SWIR2/SWIR1,
         NDVI   = ((NIR-RED)/(NIR+RED)))

ggplot(reference_spectra, aes(NDVI, SWIR_ratio, color = cover)) +
  geom_point() +
  scale_colour_viridis_d(option = "D") +
  theme_minimal() 
```

## plot complete data

```{r, cache=TRUE}
# No dissagregation
ggplot(fs_LND, aes(NDVI, SWIR_ratio)) +
  geom_bin2d(bins = 300) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2))
```

Here we can see that there are two dense zones in the spectral feature spaces. The smaller and fainter zone is predominately populated by landsat 7 pixels (clearly seen in the following plot)

## facet by sensor
```{r, cache=TRUE}
ggplot(fs_LND, aes(NDVI, SWIR_ratio)) +
  geom_bin2d(bins = 200) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~sensor)
```

Here we can see that the spectral library fits particularly well to Landsat 8 (probably do to the fact that is was created using lansat 8 pixels?). 

## facet by month
```{r, cache=TRUE}

ggplot(fs_LND, aes(NDVI, SWIR_ratio)) +
  geom_bin2d(bins = 100) +
  geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
  scale_fill_continuous(type = "viridis") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
  facet_wrap(~month)
```

Here we can also see that across all sensors, the spectral library seems to triangulate the data points better during growing season months. 

```{r, cache=TRUE, include==FALSE}

# old computationally exorbinant method of and plotting density plots 

# # Get density of points in 2 dimensions.
# # @param x A numeric vector.
# # @param y A numeric vector.
# # @param n Create a square n by n grid to compute density.
# # @return The density within each square.
# get_density <- function(x, y, ...) {
#   dens <- MASS::kde2d(x, y, ...)
#   ix <- findInterval(x, dens$x)
#   iy <- findInterval(y, dens$y)
#   ii <- cbind(ix, iy)
#   return(dens$z[ii])
# }
# 
# fs_density <- as.data.frame(fs_LND) %>%
#   filter(SWIR_ratio != "Inf") %>% 
#   mutate(density = get_density((.)$NDVI, (.)$SWIR_ratio, n = 500)) 

# ggplot(fs_density, aes(NDVI, SWIR_ratio, color = density)) +
#   geom_point() +
#   geom_point(data=reference_spectra, aes(NDVI, SWIR_ratio), color ="red") +
#   scale_colour_viridis_c(option = "D") +
#   theme_minimal() +
#   scale_x_continuous(expand = c(0, 0), limits = c(-0.2, 1)) +
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2)) +
#   facet_wrap(~sensor)

```

# single location band compairison

## Overview of observations over time by sensor

```{r, cache=TRUE}

 fs_LND %>% 
  group_by(year, sensor) %>% 
  summarise(n_observation = n()) %>% 

ggplot(., aes(year, n_observation, color=sensor)) +
  geom_line() +
  scale_fill_continuous(type = "viridis") +
  scale_colour_viridis_d(option = "D") +
  theme_minimal() 

```

##  *Exact* matches of location and *loose* match time of observation
```{r, cache=TRUE, cache=TRUE}
fs_LND_loose <- fs_LND_long %>% 
  filter(sensor %in% c("LND07", "LND08")) %>% 
  

  mutate(obs_time_place = paste0(plyr::round_any(doy, 2, f = floor),"_", year,"_",POINT_ID)) %>% 
  group_by(obs_time_place) %>% 
  filter(n() > 6)


```

Total number of observations that match the loose condition of being within ten days of each other: `r (nrow(fs_LND_loose)/6) `

Days of year with available observations: `r sort(unique(fs_LND_loose$doy)) `



```{r, cache=TRUE}
for (i in 1:10) {
  
  scens <- unique(fs_LND_loose$obs_time_place)[c((((i-1)*20)+1):((i)*20))]
  
  fs_LND_loose_subset <- fs_LND_loose %>% 
    filter(obs_time_place %in% scens) 
  
  print(ggplot( fs_LND_loose_subset, aes(wavelength_num, reflectance, color=sensor, fill=sensor)) +
    # geom_density(alpha = 0.05) +
    geom_line() +
   # scale_colour_viridis_d(option = "D") +
    #scale_fill_viridis_d(option = "D") +
    scale_x_continuous(expand = c(0, 0)) +
    #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
    theme_minimal() +
    guides(col = guide_legend(nrow = 3))+
     theme(legend.position = "bottom", 
           axis.text.y = element_text(angle = 45),
           strip.text.y = element_text(size = 8, angle = 330))  +
    facet_wrap( ~obs_time_place) )
  }

```


Summary statistics for sensor variance between all sensors
```{r, cache=TRUE}

fs_LND_sensor_summary <- fs_LND_long %>% 
  mutate(obs_time_place = paste0( round(doy, digits = -1),"_", year,"_",POINT_ID)) %>% 
  group_by(obs_time_place) %>% 
  filter(n() > 6) %>% 
  group_by(obs_time_place, sensor) %>% 
  summarise(SD = sd(reflectance))

ggplot(fs_LND_sensor_summary, aes(x = SD, y = sensor, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "SD", option = "D") +
 # scale_x_continuous(expand = c(0, 0), limits = c(0, 15000)) +
  theme_minimal()

```

### Not fully developed, But here are some additional plots with exact location matches/

## compairison of LND 7 and 8

doy agnostic plots for single locations

```{r, cache=TRUE, warning=FALSE}

fs_LND_7_8 <- fs_LND_long %>% 
  filter(sensor %in% c("LND07", "LND08"),
         year == 2020, 
         doy %in% c(93:120),
         POINT_ID %in% c(40763060:40643065)) 


fs_LND_7_8 <- fs_LND_long %>% 
  filter(sensor %in% c("LND07", "LND08"),
         year == 2020, 
         doy %in% c(96:319),
         POINT_ID %in% c(40763060:40764068)) 


# facet wrap by wavelength
ggplot(fs_LND_7_8, aes(reflectance, color=wavelength, fill=wavelength)) +
  geom_density(alpha = 0.05) +
  #geom_jitter() +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
  theme_minimal() +
  guides(col = guide_legend(nrow = 3))+
  theme(legend.position = "bottom")  +
   facet_grid(rows = vars(POINT_ID), cols = vars(sensor))


```

## *Exact* matches of location and time of observation
```{r, cache=TRUE, cache=TRUE, warning=FALSE}


fs_LND_exact <- fs_LND_long %>% 
  mutate(obs_time_place = paste0(doy,"_", year,"_",POINT_ID)) %>% 
  group_by(obs_time_place) %>% 
  filter(n() > 6)

unique(fs_LND_exact$sensor)


# facet wrap by wavelength
ggplot(fs_LND_exact[c(8000:15000),], aes(reflectance, color=wavelength, fill=wavelength)) +
  geom_density(alpha = 0.05) +
  #geom_jitter() +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.02)) +
  theme_minimal() +
  guides(col = guide_legend(nrow = 3))+
  theme(legend.position = "bottom")  +
   facet_grid(rows = vars(doy), cols = vars(sensor)) 

```

Only observation for landsat 7 and 9 exactly match. Visually the correspondence between the spectra is pretty decent though.


